---
title: "Calculating-Forage-Production"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating-Forage-Production}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(grazer)
```

# Introduction

This vignette demonstrates how to calculate forage production/biomass using the `grazer` package.

Forage production is the amount of forage available for grazing animals. It is usually expressed in terms of dry matter (DM) per unit area (e.g., kg/ha, lb/acre). Forage production is a key component of grazing management, as it determines the amount of forage available for grazing animals.

# Calculating Forage Production

First, we will load some example data to demonstrate how to calculate forage production using the `grazer` package.

```{r load example data}
file_path <- fs::path_package(
  "extdata/example_forage_data.csv",
  package = "grazer"
)

forage_data <- readr::read_csv(file_path)
```
