---
title: "Calculating Forage Production"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating forage production}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to calculate forage production/biomass using the `grazer` package.

Forage production is the amount of forage available for grazing animals. It is usually expressed in terms of dry matter (DM) per unit area (e.g., kg/ha, lb/acre). Forage production is a key component of grazing management, as it determines the amount of forage available for grazing animals.

Forage production data is usually collected using a quadrat or a sampling frame. A quadrat is a square or rectangular frame that is placed on the ground to sample vegetation. The size of the quadrat may vary depending on the study design and the vegetation type. **Note that we do not currently support the use of circular sampling frames, sometimes called "hoops".**

Vegetation is clipped from within the sampling frame, dried, and weighed to determine the dry matter (DM) production per unit area.

The forage production is then calculated by averaging the dry matter production from all sampling locations within a pasture and then scaling it to the desired unit (e.g., kg/ha, lb/acre).

## Calculating Forage Production

First, we will load some example data to demonstrate how to calculate forage production using the `grazer` package.

Load the required packages and the example data:

```{r setup}
library(grazer)
library(dplyr)
library(fs)
library(readr)
```

```{r load example data}
file_path <- fs::path_package(
  "extdata/example_forage_data.csv",
  package = "grazer"
)

forage_data <- readr::read_csv(file_path)
```

Each row represents a different pasture, and the columns (e.g., Sample_1, Sample_2) represents the weight of clipped and dried forage (in grams) within the sampling frame at multiple sample sites within the same pasture.

```{r head of forage data}
head(forage_data)
```

The area covered by a sampling frame is not typically recorded in the data, so we need to specify the dimensions of the sampling frame to calculate the area covered by the sampling frame. In this example, we will assume that the sampling frame is 50 cm x 50 cm (0.25 m^2).

```{r calculate plot area}
sample_area <- grazer::calc_plot_area(
  dim_x = 50,
  dim_y = 50,
  units = "cm"
)
```

Next, we will calculate the forage production per unit area (kg/ha) for each pasture by averaging the forage production across all sampling locations within each pasture and scaling it to the desired unit.

First, we need to calculate an average forage production value for each pasture. The average forage production is calculated by taking the mean of the forage production values from all sampling locations within a pasture.

```{r calculate average forage production using base R}
forage_data$Avg <- rowMeans(forage_data[, ], na.rm = TRUE)

head(forage_data)
```

Alternatively, you can use `dplyr` to calculate the average forage production value for each pasture if you prefer the dplyr syntax.

```{r calculate average forage production using dplyr}
forage_data <- dplyr::mutate(
  forage_data,
  Avg = rowMeans(across(starts_with("Sample")), na.rm = TRUE)
)
head(forage_data)
```

Finally, we can calculate the forage production per unit area (kg/ha) for each pasture using the `calc_forage_prod` function from the `grazer` package.

Note that we chose our output units to be "kg/ha" in this example. You can choose other units such as "lb/acre" by changing the `units` argument in the `calc_forage_prod` function.

```{r calculate forage production kg/ha}
forage_production <- grazer::calc_forage_prod(
  sample_wt = forage_data$Avg,
  sample_units = "g",
  sample_area = sample_area,
  output_units = "kg/ha"
)
head(forage_production)
```

Here's the same calculation, but with the output units set to "lb/acre".

```{r calculate forage production lb/acre}
forage_production <- grazer::calc_forage_prod(
  sample_wt = forage_data$Avg,
  sample_units = "g",
  sample_area = sample_area,
  output_units = "lb/acre"
)
head(forage_production)
```

Alternatively, we can use the `dplyr` package to create a new column in the `forage_data` data frame that contains the forage production per unit area (kg/ha) for each pasture.

The biggest different between the two methods is that the `dplyr` method will keep the original data frame structure, while the `grazer` method will return a vector of the same length as the input data.

```{r calculate forage production using dplyr}
forage_production <- dplyr::mutate(
  forage_data,
  kg_ha = grazer::calc_forage_prod(
    sample_wt = Avg,
    sample_units = "g",
    sample_area = sample_area,
    output_units = "kg/ha"
  )
)
head(forage_production)
```

Here's the same calculation, but with the output units set to "lb/acre".

```{r calculate forage production in lb/acre}
forage_production <- dplyr::mutate(
  forage_data,
  lb_acre = grazer::calc_forage_prod(
    sample_wt = Avg,
    sample_units = "g",
    sample_area = sample_area,
    output_units = "lb/acre"
  )
)
head(forage_production)
```

## Conclusion

The `grazer` package provides a convenient way to calculate forage production per unit area using data collected from sampling frames. This information is essential for grazing management, as it helps determine the amount of forage available for grazing animals.

I hope this vignette has been helpful in demonstrating how to calculate forage production using the `grazer` package. If you have any questions or feedback, please feel free to reach out to the package authors/maintainers.
